Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-23 paper.qmd
Add function to generate axes

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.45, -0.3)) +
  ylim(c(-0.05, 0.2)) +
  interior_annotation("c", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

five_gau_proj_umap_model2 <- projected_df |>
  filter(type == "data") |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  xlim(c(-0.13, 0.13)) +
  ylim(c(-0.13, 0.13)) +
  interior_annotation("d", 
                      position = c(0.08, 0.9),
                      cex = 3)

>>>>

```{r}
#| label: gen-axes

gen_axes <- function(data, proj_data, limits = 1) {

  rng <- range(proj_data)
  proj_data <- proj_data/max(abs(rng))
  colnames(proj_data) <- c("P1", "P2")
  proj_data <- data.frame(proj_data)
  obs_labels <- as.character(1:nrow(data))

  axis_labels <- colnames(data)
  rownames(axes) <- axis_labels

  theta <- seq(0, 2 * pi, length = 50)
  circle <- data.frame(c1 = adj(cos(theta)), 
                       c2 = adj(sin(theta)))

  return(list(axes = axes, circle = circle))

}
```

@@

axes_obj <- gen_axes(
  data = scaled_gau_data, 
  proj_data = projected_df, 
  limits = 1)

axes <- axes_obj$axes
circle <- axes_obj$circle

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.45, -0.3)) +
  ylim(c(-0.05, 0.2)) +
  interior_annotation("c", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

axes_obj <- gen_axes(
  data = scaled_gau_data, 
  proj_data = projected_df, 
  limits = 1)

axes <- axes_obj$axes
circle <- axes_obj$circle

five_gau_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.13, 0.13)) +
  ylim(c(-0.13, 0.13)) +
  interior_annotation("d", 
                      position = c(0.08, 0.9),
                      cex = 3)
